# Exploratory Analysis

```{r, echo=FALSE}
library(Amelia)
library(dplyr)
library(ggplot2)
library(caret)
library(doMC)

registerDoMC(cores = 8)
```

## Lets read raw data

```{r read_files}
train_file <- "pml-training.csv"
test_file <- "pml-testing.csv"

train_raw <- read.csv(train_file, na.strings=c("NA", "", "#DIV/0!"))
test_raw <- read.csv(test_file, na.strings=c("NA", "", "#DIV/0!"))
```

What is missing:
```{r}
missmap(train_raw)
```

#Data Cleaning

We need to remove features with many NAs.

```{r cleaning_data}
all_features <- names(train_raw)
num_train_cases <- nrow(train_raw)

mostly_na_features <- sapply(all_features, function(x) {
    sum(is.na(train_raw[, x])) > num_train_cases*0.9
})

non_na_features <- all_features[ ! mostly_na_features]
features <- setdiff(non_na_features, c("X", "classe"))

train_no_na <- train_raw[ , c(features, "classe")]
test_no_na <- test_raw[ , c(features, "problem_id")]
dim(train_no_na)
dim(test_no_na)
```

# Model Building

Lets split data into test(to estimate out of sample error) and training dataset.

```{r model}
in_train <- createDataPartition(train_no_na$classe, p=0.7, list=FALSE) 
train_set <- train_no_na[in_train, ]
out_of_sample <- train_no_na[-in_train, ]
```

Lets start with simple random forest:
```{r out_of_sample_error}
control <- trainControl("repeatedcv", repeats=3)
rf_model <- train(classe ~ ., data=train_set, method="rf",
    trControl=control)
```

## Feature Importance
```{r}
plot(varImp(rf_model))
```


# Out of Sample Error

Lets estimate error on unseen examples, but for the cases when we know truth:

```{r unseen_prediction}
out_of_sample_prediction <- predict(rf_model, newdata=out_of_sample)
confusionMatrix(data=out_of_sample_prediction, out_of_sample$classe)
```

So the model that we build is highly accurate > 98%.

# Prediction

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

prediction <- predict(rf_model, newdata=test_no_na)
pml_write_files(prediction)
```









